Analysis of NLA17 Dissolved Gas Indicator

### 2020-02-25 scripts/dataMunge.Rmd
Updated dataMunge and dgIndicatorAnalysis with improved data set (better BP data, a few fixes to coordinates).

The dissolved gas data were currated in the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-11-12.txt.  Data on waterbody surface area and design weights were provided by Karen Blocksom on 12/13/2019.  These data files are stored in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

@@

```{r}
# Read dissolved gas data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2019-11-12.txt"),
                 header = TRUE, sep = "\t", as.is = TRUE)

@@

# Read in temp profile data
temp <- read.csv(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/NLA17_Profile_Data.csv"),
                    as.is = TRUE)
```

@@

This leaves one site in WI that is in the dissolved gas file, but not in the weights + lake area file.  The Chain of Custody form verifies receipt of this sample.  Per Karen Blocksom, this should be SITE)ID NLA17_WI-10008, visit 1.  
``` {r}
# Will merge datasets on SITE_ID field.  Are all site.id values in dg present in wt.area?
dg %>% filter(!(site.id %in% wt.area$SITE_ID)) %>% distinct(site.id, visit.no) # Just a single WI site

# Correct site code
dg <- dg %>%
  mutate(site.id = replace(site.id, site.id == "NLA17_WI-10018" & visit.no == 1, "NLA17_WI-10008"))
```

@@

``` {r}
# Any site id values in wt.area, but not present in dg?
wt.area %>% filter(!(SITE_ID %in% dg$site.id)) %>% select(SITE_ID, EVAL_CAT, SITETYPE) # h23 sites w/out dissolved gas data
```

@@

## Calculate stratification indices from temperature profile data
Water temperature profiles were measured at the index site in each waterbody.  A variety of metrics describing the degree of thermal stratification can be extracted from these data.  Webb et al. used the Brunt–Väisälä buoyancy frequency as an indicator of thermal stratification strength.  We will do the same here.

Three unique site visits contain only NA for temperature.  Two of the incidents include notes that sonde was forgotten or malfunctioning. Two of the three sites were revisited later in the study and temperature data were collected.  Lets assign data from the second site visit to the first visit when no temperature data were collected.  This leaves only NLA17_CA-10082, visit 1, without temperature data.

```{r}
# Sites that only contain NAs.
missingTemp <- temp %>% group_by(SITE_ID, VISIT_NO) %>% 
  summarise(all.na = all(is.na(TEMPERATURE))) %>% 
  filter(all.na == TRUE) %>%
  select(SITE_ID, VISIT_NO)

# View sites with missing temp.
missingTemp

# Two of the sites have data from second site visit
temp %>% filter(SITE_ID %in% missingTemp$SITE_ID) %>%
  select(SITE_ID, VISIT_NO, TEMPERATURE)

# Extract data from second visit to assign to first visit
replacementData <- temp %>% filter(SITE_ID %in% missingTemp$SITE_ID, VISIT_NO == 2) %>%
  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE) %>%
  mutate(VISIT_NO = 1) # change visit no to one for join

# Join the data
temp1 <- full_join(temp, replacementData)

# Inspect results of join.  Looks good.  Only one site missing temperature data.
temp1 %>% filter(SITE_ID %in% missingTemp$SITE_ID) %>%
  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE) %>% 
  arrange(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE)

@@

The field crews measured temperature starting near the water surface, then at increasing depths.  A second measurement was made at the water surface before the sonde was removed from the water, therefore the file contains duplicate measurements at the shallowest depths. Duplicate measurements corrupt the buoyancy frequency function and will therefore be aggregated. 
``` {r}
# Aggregate across SITE_ID, VISIT_NO, and DEPTH.  I don't like the dplyr solution for this (group_map) and I'm not familiar with the purr::map_ functions.  I'm going old-school and busting out the trusty plyr::ddply.
temp2 <- temp1 %>% select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE) %>%
  # exclude NA, NaN, Inf...
  filter(is.finite(TEMPERATURE)) %>% # This completely removes the CA site with only NA for temp data.
  plyr::ddply(., .(SITE_ID, VISIT_NO, DEPTH), summarize, 
              TEMPERATURE.M = mean(TEMPERATURE, na.rm = TRUE))
dim(temp1) # 11003
dim(temp2) # 10071, nearly 1000 observations aggregated or stripped out due to NA
```

@@

``` {r}
# The buoyancy frequency is calculated at multiple depths, but at fewer depths than the where 
# temperature was measured.  The output of the buoyancy frequency function is shorter 
# than the input, which is incompatible with dplyr mutate and summarize.  The dplyr alternative
# is group_map which produces a list, but the grouping variables are stripped from the list
# elements.  I really don't like this.  I think this can be addressed by combining group_map
# with purr::map_ functions, but I'm not familiar with purr.  Rather, I'll use the old school
# split -lapply - do(rbind) approach.
bf <- temp2 %>%
  group_split(SITE_ID, VISIT_NO) %>% # split into list of df
  lapply(function(x) {
    data.frame(BF = buoyancy.freq(wtr = x$TEMPERATURE.M, depths = x$DEPTH), # calculate buoyance frequency
               SITE_ID = unique(x$SITE_ID), # add site ID
               VISIT_NO = unique(x$VISIT_NO), # add visit no
               stringsAsFactors = FALSE)}) %>% 
  lapply(function(x) x %>% mutate(DEPTHS.BF = attr(BF, "depths"))) %>% # depths stored as attribute, assign to column
  do.call("rbind", .) # collapse to df

@@

# These site have a single temp measurement made at shallow depths (<1m).  Presumably these are 
# shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0.
bf %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT) 
temp2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT)

@@

>>>>

The dissolved gas data were currated in the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-02-24.txt.  Data on waterbody surface area and design weights were provided by Karen Blocksom on 12/13/2019.  These data files are stored in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

@@

```{r}
# Read dissolved gas data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2020-02-24.txt"),
                 header = TRUE, sep = "\t", as.is = TRUE)

@@

# Read in temp profile data
temp.o2 <- read.csv(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/NLA17_Profile_Data_Updated_11Feb2020.csv"),
                    as.is = TRUE)
```

@@

@@

``` {r}
# Any site id values in wt.area, but not present in dg?
wt.area %>% filter(!(SITE_ID %in% dg$site.id)) %>% select(SITE_ID, EVAL_CAT, SITETYPE) # 23 sites w/out dissolved gas data
```

@@

## Calculate dissolved oxygen and stratification indices from profile data
Water temperature and dissolved oxygen profiles were measured at the index site in each waterbody.  A variety of metrics describing the degree of thermal stratification can be extracted from these data.  Webb et al. used the Brunt–Väisälä buoyancy frequency as an indicator of thermal stratification strength.  We will do the same here.

Three site visits contain only NA for temperature.  Four site visits contain only NA for do.  Two of the incidents include notes that sonde was forgotten or malfunctioning. Two of the three sites were revisited later in the study and temperature data were collected.  Lets assign data from the second site visit to the first visit when no temperature data were collected.  This leaves only NLA17_CA-10082, visit 1, without temperature data.  NLA17_ID-10084 has no DO data and was not resampled.  Data sheet comment "Probe not measuring LDO for all depths".

```{r}
# Sites that only contain NAs.
missing <- temp.o2 %>% group_by(SITE_ID, VISIT_NO) %>% 
  summarise(all.temp = all(is.na(TEMPERATURE)),
            all.do = all(is.na(OXYGEN))) %>% 
  filter(all.temp == TRUE | all.do == TRUE) %>%
  select(SITE_ID, VISIT_NO, all.temp, all.do)

# View sites with missing temp/do
missing

# Two of the sites have data from second site visit
temp.o2 %>% filter(SITE_ID %in% missing$SITE_ID) %>%
  select(SITE_ID, VISIT_NO, TEMPERATURE, OXYGEN)

# Extract data from second visit to assign to first visit
replacementData <- temp.o2 %>% filter(SITE_ID %in% missing$SITE_ID, VISIT_NO == 2) %>%
  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>%
  mutate(VISIT_NO = 1) # change visit no to one for join

# Join the data
temp.o21 <- full_join(temp.o2, replacementData)

# Inspect results of join.  Looks good.  
# Only one site missing temperature and DO data. (NLA17_CA-10082)
# One site missing all DO data (NLA17_ID-10084)
temp.o21 %>% filter(SITE_ID %in% missing$SITE_ID) %>%
  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>% 
  arrange(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN)

@@

The field crews measured temperature/DO starting near the water surface, then at increasing depths.  A second measurement was made at the water surface before the sonde was removed from the water, therefore the file contains duplicate measurements at the shallowest depths. Duplicate measurements corrupt the buoyancy frequency function and will therefore be aggregated. 
``` {r}
# Aggregate across SITE_ID, VISIT_NO, and DEPTH.  I don't like the dplyr solution for this (group_map) and I'm not familiar with the purr::map_ functions.  I'm going old-school and busting out the trusty plyr::ddply.
temp.o2.2 <- temp.o21 %>% select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>%
  plyr::ddply(., .(SITE_ID, VISIT_NO, DEPTH), summarize, 
              TEMPERATURE.M = mean(TEMPERATURE, na.rm = TRUE),
              OXYGEN.M = mean(OXYGEN, na.rm = TRUE))

dim(temp.o21) # 11003
dim(temp.o2.2) # 10133, 870 observations aggregated
```

@@

``` {r}
# The buoyancy frequency is calculated at multiple depths, but at fewer depths than the where 
# temperature was measured.  The output of the buoyancy frequency function is shorter 
# than the input, which is incompatible with dplyr mutate and summarize.  The dplyr alternative
# is group_map which produces a list, but the grouping variables are stripped from the list
# elements.  I really don't like this.  I think this can be addressed by combining group_map
# with purr::map_ functions, but I'm not familiar with purr.  Rather, I'll use the old school
# split -lapply - do(rbind) approach.

# First, strip out any NA, NaN, Inf...etc

bf <- temp.o2.2 %>%
  filter(is.finite(TEMPERATURE.M)) %>% # This completely removes the CA site with only NA for temp data.
  group_split(SITE_ID, VISIT_NO) %>% # split into list of df
  lapply(function(x) {
    data.frame(BF = buoyancy.freq(wtr = x$TEMPERATURE.M, depths = x$DEPTH), # calculate buoyance frequency
               SITE_ID = unique(x$SITE_ID), # add site ID
               VISIT_NO = unique(x$VISIT_NO), # add visit no
               stringsAsFactors = FALSE)}) %>% 
  lapply(function(x) x %>% mutate(DEPTHS.BF = attr(BF, "depths"))) %>% # depths stored as attribute, assign to column
  do.call("rbind", .) # collapse to df

@@

# These site have a single temp measurement made at shallow depths (<1m).  Presumably these are 
# shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0 (see below)
bf %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT) 
temp.o2.2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT)

@@

Webb et al. used the 'surface' and 'bottom' water DO percent saturation and temperature as predictors, but didn't give more details on how these indices were calculated.  We will do something similar and define the 'shallowest' and 'deepest' measurements as 'surface' and 'bottom', respectively, but will use mg.l for dissolved oxygen.
```{r}
# o2 surface (s) and bottom (b)
o2.s.b <- 
temp.o2.2 %>% 
  group_by(SITE_ID, VISIT_NO) %>%
  slice(c(which.min(DEPTH), which.max(DEPTH))) %>% # min and max depth by group
  arrange(SITE_ID, VISIT_NO, DEPTH) %>%
    mutate(DEPTH = ifelse(row_number() == 1,
                        "surface",
                        "bottom")) %>%
  pivot_wider(id_cols = c(SITE_ID, VISIT_NO), 
              names_from = DEPTH, 
              values_from = c(TEMPERATURE.M, OXYGEN.M)) %>%
  rename(temp.surf = TEMPERATURE.M_surface,
         temp.bottom = TEMPERATURE.M_bottom,
         o2.surf = OXYGEN.M_surface,
         o2.bottom = OXYGEN.M_bottom)

o2.s.b

These calculations produced 8 site visits with at least one NA for surface/bottom temp/DO.

```{r}
o2.s.b %>% filter(is.na(temp.surf) | is.na(temp.bottom) | is.na(o2.surf) | is.na(temp.bottom))
```

o2.surf at NLA17_FL-10029 is a case with no DO at shallowest depth, but good measurements below.  Replace with next shallowest value.
``` {r}
# Inspect raw data
temp.o2.2 %>% filter(SITE_ID == "NLA17_FL-10029", VISIT_NO == 1)

# update data
o2.s.b <- o2.s.b %>%
  mutate(o2.surf = replace(o2.surf, 
                           SITE_ID == "NLA17_FL-10029" & VISIT_NO == 1,
                           temp.o2 %>% filter(SITE_ID=="NLA17_FL-10029",
                                              VISIT_NO == 1,
                                              DEPTH == 1) %>%
                             select(OXYGEN) %>% pull()))

```
No bottom water DO or temp at NLA17_ID-10001, visit 1 due to probe malfunction.

```{r}
temp.o2.2 %>% filter(SITE_ID == "NLA17_ID-10001", VISIT_NO == 1)
```

No DO NLA17_ID-10084, visit 1 due to probe malfunction.

```{r}
temp.o2.2 %>% filter(SITE_ID == "NLA17_ID-10084", VISIT_NO == 1)
```

No deep DO or temp at NLA17_MI-10073, visit 1 becuase no measurement taken at deepst recorded depth.  Replace with next shallowest measurement.

```{r}
# look at raw data
temp.o2.2 %>% filter(SITE_ID == "NLA17_MI-10073", VISIT_NO == 1)

# update aggregated data
o2.s.b <- o2.s.b %>%
  mutate(o2.bottom = replace(o2.bottom, 
                           SITE_ID == "NLA17_MI-10073" & VISIT_NO == 1,
                           temp.o2.2 %>% filter(SITE_ID=="NLA17_MI-10073",
                                              VISIT_NO == 1,
                                              DEPTH == 1.3) %>%
                             select(OXYGEN.M) %>% pull()),
         temp.bottom = replace(temp.bottom, 
                           SITE_ID == "NLA17_MI-10073" & VISIT_NO == 1,
                           temp.o2.2 %>% filter(SITE_ID=="NLA17_MI-10073",
                                              VISIT_NO == 1,
                                              DEPTH == 1.3) %>%
                             select(TEMPERATURE.M) %>% pull()))
```

No surface temp at NLA17_NE-10060, visit 1 becuase no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)

```{r}
# look at raw data
temp.o2.2 %>% filter(SITE_ID == "NLA17_NE-10060", VISIT_NO == 1)

o2.s.b <- o2.s.b %>%
  mutate(temp.surf = replace(temp.surf, 
                           SITE_ID == "NLA17_NE-10060" & VISIT_NO == 1,
                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NE-10060",
                                              VISIT_NO == 1,
                                              DEPTH == 0.1) %>%
                             select(TEMPERATURE.M) %>% pull()))
```

No bottom temp or O2 at NLA17_NH-10063, visit 1 becuase no measurement taken at greatest depth (8m). Replace with next deeptest measurement (7.5m)

```{r}
# look at raw data
temp.o2.2 %>% filter(SITE_ID == "NLA17_NH-10063", VISIT_NO == 1)

o2.s.b <- o2.s.b %>%
  mutate(temp.bottom = replace(temp.bottom, 
                           SITE_ID == "NLA17_NH-10063" & VISIT_NO == 1,
                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NH-10063",
                                              VISIT_NO == 1,
                                              DEPTH == 7.5) %>%
                             select(TEMPERATURE.M) %>% pull()),
         o2.bottom = replace(o2.bottom, 
                           SITE_ID == "NLA17_NH-10063" & VISIT_NO == 1,
                           temp.o2.2 %>% filter(SITE_ID=="NLA17_NH-10063",
                                              VISIT_NO == 1,
                                              DEPTH == 7.5) %>%
                             select(OXYGEN.M) %>% pull()))
```

No surface temp or O2 at NLA17_OR-10003, visit 1 becuase no measurement taken at shallowest depth (0m). Replace with next shallowest measurement (0.1m)

```{r}
# look at raw data
temp.o2.2 %>% filter(SITE_ID == "NLA17_OR-10003", VISIT_NO == 1)

o2.s.b <- o2.s.b %>%
  mutate(temp.surf = replace(temp.surf, 
                           SITE_ID == "NLA17_OR-10003" & VISIT_NO == 1,
                           temp.o2.2 %>% filter(SITE_ID=="NLA17_OR-10003",
                                              VISIT_NO == 1,
                                              DEPTH == 0.1) %>%
                             select(TEMPERATURE.M) %>% pull()),
         o2.surf = replace(o2.surf, 
                           SITE_ID == "NLA17_OR-10003" & VISIT_NO == 1,
                           temp.o2.2 %>% filter(SITE_ID=="NLA17_OR-10003",
                                              VISIT_NO == 1,
                                              DEPTH == 0.1) %>%
                             select(OXYGEN.M) %>% pull()))
```

No bottom temp or O2 at NLA17_SD-10057, visit 1 becuase no measurement taken at deepest depth (4m). Replace with next deepest measurement (3m)

```{r}
# look at raw data
temp.o2.2 %>% filter(SITE_ID == "NLA17_SD-10057", VISIT_NO == 1)

o2.s.b <- o2.s.b %>%
  mutate(temp.bottom = replace(temp.bottom, 
                           SITE_ID == "NLA17_SD-10057" & VISIT_NO == 1,
                           temp.o2.2 %>% filter(SITE_ID=="NLA17_SD-10057",
                                              VISIT_NO == 1,
                                              DEPTH == 3) %>%
                             select(TEMPERATURE.M) %>% pull()),
         o2.bottom = replace(o2.bottom, 
                           SITE_ID == "NLA17_SD-10057" & VISIT_NO == 1,
                           temp.o2.2 %>% filter(SITE_ID=="NLA17_SD-10057",
                                              VISIT_NO == 1,
                                              DEPTH == 3) %>%
                             select(OXYGEN.M) %>% pull()))
```

Only remainining missing values are observations where field crews noted sonde malfunctions.

```{r}
o2.s.b %>% filter(is.na(temp.surf) | is.na(temp.bottom) | is.na(o2.surf) | is.na(temp.bottom))
```

Merge bottom and surface DO and temperature with dg data.

```{r}
# are all SITE_ID values in o2 and temp data also present in dg?
o2.s.b %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007
print(n= Inf)

# are all site.id values in dg also present in o2 and temp data?
# Just CA site, as expected.
dg %>% filter(!(site.id %in% o2.s.b$SITE_ID)) %>% # no
  select(site.id, visit.no)

# dim(o2.s.b) # 1209, 6
# dim(dg) # 2369, 375

# Left join: retain all dg values, exclude non-matching bf values (i.e. OR and AZ)
dg <- left_join(dg, o2.s.b, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))  
# dim(dg) # 2369, 379, good!
```

### 2020-02-25 scripts/dgIndicatorAnalysis.Rmd
Updated dataMunge and dgIndicatorAnalysis with improved data set (better BP data, a few fixes to coordinates).

```{r results='hide', message=FALSE, warning=FALSE}
# load libraries
library(sf) # spatial data
library(gstat) # lagged scatterplot, variogram
#library(spdplyr) # apply dplyr verbs to sp objects.  gstat functions require sp objects
library(tidyverse) # dplyr, ggplot
library(janitor) # clean names
library(spdep) # k-NN and Moran plot

@@

## Getting data and filtering
The dissolved gas data were currated the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-10-25.txt.  A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  This file was futher manipulated in "C:\Users\JBEAULIE\OneDrive - Environmental Protection Agency (EPA)\gitRepository\DissolvedGasNla\scripts\dataMunge.Rmd" and written out as an .RData object which is imported below.

```{r}
# Read data file
load(paste0(localPath, 
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.RData"))

@@

# extract population estimates for ecoregion and nation
pop.eco <- pop %>% 
  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
         statistic == "50Pct") # extract mean estimate

@@

# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "Aggr_Ecoregions9_2015")

@@

# Check CRS
st_crs(ecoR) # no EPSG code
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
st_crs(ecoR) # 5070
```

@@

We can also look at population estimates of N2O saturation ratio by ecoregion.  Wow, the mean N2O saturation is less than one for all ecoregions and the nation!  95% CI overlap one in only 3 levels of the stratification.  This is pretty big news. 
```{r}

@@

## Geopackage
Create a simple feature on entire dg data frame, then split out a simple feature, df_sf2, by filtering sample_source = DG.  The geodatabase will contain two simple features, the dissolved gas sampled points/lakes and the nine aggregated ecoregions.
```{r, geopackage}
library(janitor) # clean_names function
names(dg)

@@

ecoR1 <- st_transform(ecoR, 4326)
st_crs(ecoR1)

>>>>

```{r results='hide', message=FALSE, warning=FALSE}
# load libraries
library(sf) # spatial data
library(gstat) # lagged scatterplot, variogram
library(tidyverse) # dplyr, ggplot
library(janitor) # clean names
library(spdep) # k-NN and Moran plot

@@

## Getting data and filtering
The dissolved gas data were currated the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-02-24.txt.  A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  This file was futher manipulated in "C://Users//JBEAULIE//OneDrive-Environmental Protection Agency (EPA)//gitRepository//DissolvedGasNla//scripts//dataMunge.Rmd" and written out as an .RData object, which is imported below.

```{r}
# Read data file
load(paste0(localPath, 
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.2020-02-24.RData"))

# Strip out air samples
dg <- dg %>% filter(sample.source == "DG")

@@

# extract population estimates for ecoregion and nation
pop.eco <- pop %>% 
  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
         statistic == "Mean") # extract mean estimate

@@

# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple")

@@

# Check CRS
st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
st_crs(ecoR) # 5070
```

@@

We can also look at population estimates of N2O saturation ratio by ecoregion.  Wow, despite the prevalance of N2O undersaturated waterbodies, the 95% CI for the mean N2O saturation ratio encompasses 1 in all ecoregions except the Northern Plains and Northern Appalachians. 
```{r}

@@

## Geopackage
Create a simple feature on entire dg data frame, then split out a simple feature, df_sf2, by filtering sample_source = DG.  The geodatabase will contain two simple features, the dissolved gas sampled points/lakes and the nine aggregated ecoregions.
```{r, geopackage, eval=FALSE, echo=FALSE,}
library(janitor) # clean_names function
names(dg)

@@

ecoR1 <- st_transform(ecoR, 4326) %>% 
  select(OBJECTID, WSA9, WSA9_NAME) # Grab variables of interest
st_crs(ecoR1)
